# Results {#results}

These are the variables I'm using LIST/TABLE/CALCULATION

## Modeling turnout

```{r}
summary(best_linear_turnout)
ggplot(best_linear_turnout, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")

ggplot(best_linear_turnout, aes(x = .resid)) +
  geom_histogram() +
  xlab("Residuals")
# linear model is not half bad at turnout
# slightly right skewed residuals, a few super high outliers
# investigate these
```

## Modeling overvoting and undervoting

I have some results and they are here. First model is the overvoting linear model, second is undervoting linear.

```{r}
summary(best_linear_over)
summary(best_linear_under)
```

Writeup of said results is forthcoming because it's way past my bedtime right now. Not sure why the undervote model is the full model - nothing in it is significant, but apparently that had the best MSE. I'll check it more seriously tomorrow.

And these are logistic for over and under - still need to do training/test or cross validation, plus better model selection.

```{r}
summary(logit_over)
summary(logit_under)
```

